### 2017-10-16 mcmc.tex
acronym improvements; abstract tweak

% to-do
% -----
% - deal with \MCMC\ macro in footnotes.
% - Deal with all mentions of ``DFM'' or ``HOGG''
% - Go through github issues, if any.
% - Need to mention case where there is an MC integral INSIDE the likelihood function
% - Need to mention testing the likelihood function, by, say, 1-d scans back and forth.
% - Kill all issues in this list, above, and then:
%   - Post on arXiv
%   - Submit to ApJ special issue
%   - Blog about it

@@

\begin{abstract}
Markov Chain Monte Carlo (\MCMC) methods for sampling probability
density functions (combined with abundant computational resources)
have transformed the sciences, especially in performing probabilistic
inferences, or fitting models to data.
In this primarily pedagogical contribution,
we give a brief overview of the most basic \MCMC\ method and some
practical advice for the use of \MCMC\ in real inference problems.
We give advice on method choice, tuning for performance,
methods for initialization, tests of convergence, troubleshooting, and use of the chain
output to produce or report parameter estimates with associated uncertainties.
We emphasize that sampling is a method for doing integrals; this guides our thinking
about how \MCMC\ output is best used.
\end{abstract}

@@

More formally, as discussed in \sectionname~\ref{sec:sampling}, the goal of
MCMC sampling is to compute integrals of the form given in
\equationname~\ref{eq:the-real-integral} using the Monte Carlo approximation
in \equationname~\ref{eq:the-real-samples}.
The Monte Carlo error introduced by this approximation is proportional to
$\sqrt{\tau_\mathrm{int}/N}$ where $\tau_\mathrm{int}$ is the
integrated autocorrelation time and $N$ is the total number of
samples of $p(\pars)$.
In other words, $\tau_\mathrm{int}$ is the number of steps required for the
chain to produce an \emph{independent} sample.

@@

We won't go into details about how to estimate $\tau_\mathrm{int}$, but it is
notoriously difficult: It is a two-point statistic, and two-point statistics
are much harder to estimate than one-point statistics.\note{See: All of
cosmology!}
If you are not a hard-core user of MCMC, and if all you want is
heuristic indicators of convergence, then what you should take from this
section is that the autocorrelation time is involved in variance estimates,
but that it is hard to estimate.
If you want to think about autocorrelation estimation, more detailed
references can be found elsewhere.\note{%
    A canonical reference is a set of lecture notes by Alan Sokal
    \citep{sokal}.
    Another good reference is a blog post by one of us (DFM) that can be found
    at \url{http://dfm.io/posts/autocorr}.}

%Of course, the autocorrelation-time definition
%(\ref{eq:autocorrelationtime}) does not by itself provide an
%\emph{estimator} for the time. Estimating the auto-correlation time is
%notoriously difficult: It is a two-point statistic, and two-point statistics
%are much harder to estimate than one-point statistics.\note{See: All of cosmology!}
%If you are not a hard-core user of MCMC, and if all you want is heuristic indicators
%of convergence, then what you should take from this section is that the auto-correlation
%time is involved in variance estimates through \equationname~(\ref{eq:tauvar}) but that
%it is hard to estimate.

@@

If you feel drawn to give the best sample, instead give the parameter
values found by optimizing the posterior pdf, starting at the best
sample as an initialization (and sacrifice your probabilism\note{We
  choose not to count the number of papers out there that claim to be
  ``Bayesian'' but then deliver the optimum of a posterior PDF.  That
  sacrifices probabilism, but also all the useful things you get from
  delivering a posterior, like protection from over-fitting, and
  non-approximate uncertainty propagation.  You certainly don't need
  to be a Bayesian---you can do most of the same science as a
  frequentist---but if you are using \MCMC\ you probably are doing so
  because you want the benefits of Bayesianism.}).
This optimal-posterior parameter value is called the
  ``maximum \foreign{a posteriori}'' or ``MAP'' value for the parameters;
  it is like a maximum-likelihood value but regularized by the prior.
It is an estimator with some good (and some bad) properties, but it is not the point of \MCMC,
  and therefore outside the scope of this \documentname.

@@

Of course figures are always better than tables, at least in the
printed form of scientific publications.
There are several figures that are useful and should be made every time you
run MCMC and we'll describe a few here.

@@

\paragraph{Ensembles:}
As we discussed (\sectionname~\ref{sec:tuning}), one of the critical challenges in the use of
\MCMC\ is tuning the proposal pdf $q(\pars'\given\pars)$.
In problems in which all of the parameters (all of the components or entries of $\pars$) are somehow ``equivalent'',
or the parameters can be seen as components of a vector in a vector
space, there are \emph{ensemble methods} that make use of not just one
random-walker but instead many walkers to automatically generate a properly tuned
proposal distribution.
These methods make use of the distribution of a set of independent
walkers in the parameter space ($\pars$-space) to gauge the typical step-sizes and
directions at which new proposals should be made, obviating or greatly
reducing the tuning requirements.
We are co-authors on one of these methods,
\project{emcee}\note{\citet{emcee}}, which is popular in the
physical sciences, but there are others\note{A few ensemble methods that
are being used in astronomy: Differential Evolution MCMC (\citealt{demcmc,
run-dmc}) and \project{kombine} (Farr \& Farr, in preparation).}.
It is worty of note that ensemble methods have particular
peculiarities about initialization and convergence diagnostics that
are worth understanding before use.\note{For example, we find
  empirically that it is better to initialize the ensemble in a ball
  that is smaller than the full posterior pdf width, not larger.  For
  another example, the ensemble produces a set of chains that can be
  used to compute the Gelman--Rubin diagnostic \citep{gelmanrubin},
  but this is not a conservative test if the initialization was made
  in a small ball.}
Because ensemble methods are updating independent \MCMC\ chains, they
are often easy to parallelize to make use of multiple cores.

>>>>

% to-do
% -----
% - Go through github issues, if any.
% - Post on arXiv
% - Submit to ApJ special issue
% - Blog about it

@@

\begin{abstract}
Markov Chain Monte Carlo (\MCMC) methods for sampling probability
density functions (combined with abundant computational resources)
have transformed the sciences, especially in performing probabilistic
inferences, or fitting models to data.
In this primarily pedagogical contribution,
we give a brief overview of the most basic \MCMC\ method and some
practical advice for the use of \MCMC\ in real inference problems.
We give advice on method choice, tuning for performance,
methods for initialization, tests of convergence, troubleshooting, and use of the chain
output to produce or report parameter estimates with associated uncertainties.
We argue that autocorrelation time is the most important test for convergence,
as it directly connects to the variances of sampling estimates of quantities of interest.
We emphasize that sampling is a method for doing integrals; this guides our thinking
about how \MCMC\ output is best used.
\end{abstract}

@@

More formally, as discussed in \sectionname~\ref{sec:sampling}, the goal of
\MCMC\ sampling is to compute integrals of the form given in
\equationname~\ref{eq:the-real-integral} using the Monte Carlo approximation
in \equationname~\ref{eq:the-real-samples}.
The Monte Carlo error introduced by this approximation is proportional to
$\sqrt{\tau_\mathrm{int}/N}$ where $\tau_\mathrm{int}$ is the
integrated autocorrelation time and $N$ is the total number of
samples of $p(\pars)$.
In other words, $\tau_\mathrm{int}$ is the number of steps required for the
chain to produce an \emph{independent} sample.

@@

We won't go into details about how to estimate $\tau_\mathrm{int}$, but it is
notoriously difficult: It is a two-point statistic, and two-point statistics
are much harder to estimate than one-point statistics.\note{See: All of
cosmology!}
If you are not a hard-core user of \MCMC, and if all you want is
heuristic indicators of convergence, then what you should take from this
section is that the autocorrelation time is involved in variance estimates,
but that it is hard to estimate.
If you want to think about autocorrelation estimation, more detailed
references can be found elsewhere.\note{%
    A canonical reference is a set of lecture notes by Alan Sokal
    \citep{sokal}.
    Another good reference is a blog post by one of us (DFM) that can be found
    at \url{http://dfm.io/posts/autocorr}.}

%Of course, the autocorrelation-time definition
%(\ref{eq:autocorrelationtime}) does not by itself provide an
%\emph{estimator} for the time. Estimating the auto-correlation time is
%notoriously difficult: It is a two-point statistic, and two-point statistics
%are much harder to estimate than one-point statistics.\note{See: All of cosmology!}
%If you are not a hard-core user of \MCMC, and if all you want is heuristic indicators
%of convergence, then what you should take from this section is that the auto-correlation
%time is involved in variance estimates through \equationname~(\ref{eq:tauvar}) but that
%it is hard to estimate.

@@

If you feel drawn to give the best sample, instead give the parameter
values found by optimizing the posterior pdf, starting at the best
sample as an initialization (and sacrifice your probabilism\note{We
  choose not to count the number of papers out there that claim to be
  ``Bayesian'' but then deliver the optimum of a posterior PDF.  That
  sacrifices probabilism, but also all the useful things you get from
  delivering a posterior, like protection from over-fitting, and
  non-approximate uncertainty propagation.  You certainly don't need
  to be a Bayesian---you can do most of the same science as a
  frequentist---but if you are using \MCMC\ you probably are doing so
  because you want the benefits of Bayesianism.}).
This optimal-posterior parameter value is called the
  ``maximum \foreign{a posteriori}'' or ``\acronym{MAP}'' value for the parameters;
  it is like a maximum-likelihood value but regularized by the prior.
It is an estimator with some good (and some bad) properties, but it is not the point of \MCMC,
  and therefore outside the scope of this \documentname.

@@

Of course figures are always better than tables, at least in the
printed form of scientific publications.
There are several figures that are useful and should be made every time you
run \MCMC\ and we'll describe a few here.

@@

\paragraph{Ensembles:}
As we discussed (\sectionname~\ref{sec:tuning}), one of the critical challenges in the use of
\MCMC\ is tuning the proposal pdf $q(\pars'\given\pars)$.
In problems in which all of the parameters (all of the components or entries of $\pars$) are somehow ``equivalent'',
or the parameters can be seen as components of a vector in a vector
space, there are \emph{ensemble methods} that make use of not just one
random-walker but instead many walkers to automatically generate a properly tuned
proposal distribution.
These methods make use of the distribution of a set of independent
walkers in the parameter space ($\pars$-space) to gauge the typical step-sizes and
directions at which new proposals should be made, obviating or greatly
reducing the tuning requirements.
We are co-authors on one of these methods,
\project{emcee}\note{\citet{emcee}}, which is popular in the
physical sciences, but there are others\note{A few ensemble methods that
are being used in astronomy: Differential Evolution \MCMC\ (\citealt{demcmc,
run-dmc}) and \project{kombine} (Farr \& Farr, in preparation).}.
It is worty of note that ensemble methods have particular
peculiarities about initialization and convergence diagnostics that
are worth understanding before use.\note{For example, we find
  empirically that it is better to initialize the ensemble in a ball
  that is smaller than the full posterior pdf width, not larger.  For
  another example, the ensemble produces a set of chains that can be
  used to compute the Gelman--Rubin diagnostic \citep{gelmanrubin},
  but this is not a conservative test if the initialization was made
  in a small ball.}
Because ensemble methods are updating independent \MCMC\ chains, they
are often easy to parallelize to make use of multiple cores.
